\documentclass[11pt]{article}
\usepackage[a4paper, total={6in, 10in}]{geometry}
%\usepackage[hyphens]{url} % <-- new
\usepackage{natbib}

%\usepackage{graphicx} %Loading the package
%\graphicspath{{../src/sinewaves/}}


\usepackage{xspace}
\newcommand{\R}{\textsf{R}\xspace}

%% Codelinks
\usepackage{hyperref}
\usepackage{fontawesome}
\usepackage{color}
\definecolor{linkcolor}{rgb}{0.1216,0.4667,0.7059} % blue for ONLINE version
\definecolor{linkbw}{rgb}{0.0, 0.0, 0.0} % black for PRINTER version
\newcommand{\codeicon}{{\color{linkcolor}\faFileCodeO}}
\newcommand{\codecolorlink}[1]{\href{#1}{\codeicon}}

\newcommand{\codeiconbw}{{\color{linkbw}\faFileCodeO}}
\newcommand{\codebwlink}[1]{\href{#1}{\codeiconbw}}



\title{Erratum Notice} 
\author{Miguel Xochicale}
\date{\today}
%\date{30th August 2019}

\begin{document}
\maketitle
%(examples given for some common errors, edit/delete as necesary)

\begin{enumerate}
\item On all the thesis, links with the \texttt{url}:   
\texttt{https://github.com/mxochicale/phd-thesis/}
should be linked as: \texttt{https://github.com/mxochicale-phd/thesis/}.
See \cite{xochicale2019-links} for detailed changes on the \LaTeX \space files.


\item On all the thesis, the reference \texttt{xochicale2018} should be cited as
\texttt{xochicale2019}.
See \cite{xochicale2019-links} for detailed changes on the \LaTeX \space files.


\item On all the thesis, the colour of link icons were updated from black and white 
(\texttt{\codebwlink{https://github.com/mxochicale-phd/thesis/}}) to 
blue (\texttt{\codecolorlink{https://github.com/mxochicale-phd/thesis/}}).
See \cite{xochicale2019-links} for detailed changes on the \LaTeX \space files.


\item Github repository date was updated from 20 May 2019 to 30 August 2019.
See \cite{xochicale2019-links} for detailed changes in the \LaTeX \space files.


\item Submission date was updated from from 20 May 2019 to 30 August 2019.
See \cite{xochicale2019-links} for detailed changes in the \LaTeX \space files.


\item On page 113, instead of proving a link for figures of the surrogate 
data analysis, a link for a document of such figures has been added. Aditionally, 
the description for the link of the document for figures has been partially reworded. 
See \cite{xochicale2019-surrogate} for detailed changes in the \LaTeX files.

\end{enumerate}


%3. On (page), an error appeared in (description of formula etc),
%the (formula) should read:
%(corrected formula)

%4. On (page), (table/diagram/figure) contained an error in 
%(description of error), a corrected version is included below:
%(corrected table/diagram/figure etc.)

\bibliography{references}
\bibliographystyle{apalike}

\end{document}
